# Introduction

## Motivation

What questions can we ask of the wage data?
 - age, year education data
 
```{r}
library(ISLR)
library(dplyr)
library(ggplot2)
```


```{r}
Wage %>% 
  glimpse
```

```{r}
Wage %>% 
  ggplot(aes(x = age, y = wage)) + 
  geom_point() + 
  geom_smooth()
```
 
 Why do we care?
- Selfish, policy?

Other Data Examples
 - stock market
 
```{r}
ISLR::Smarket %>% 
  glimpse
```
 
```{r}
Smarket %>% 
  mutate(row_num = row_number()) %>% 
  ggplot(aes(x = Lag1, y = Today)) + 
  geom_point()
```



 - gene expression

```{r}
data.frame(
  Z1 = ISLR::NCI60$data[, 1],
  Z2 = ISLR::NCI60$data[, 2],
  label = ISLR::NCI60$labs) %>% 
  ggplot(aes(x = Z1, y = Z2, color = label)) + 
  geom_point()
```



## Where are we now?

Statistical Learning vs. 
  Machine Learning vs. 
  Statistical Analysis vs. 
  AI vs. 
  Data Analysis vs. 
  Data Science
  
I will use all of these interchangibly depending on what I've read on a given day.

## Why R?

Language focused on statistical analysis. It has some unique concepts that make it especially good at communicating statistical concepts.

Lots of people have started using it in the last 10 years

## What will this class do?

1. Provide a survey of common and powerful statistical methods.
2. Teach you how to apply them using R - including slightly easier versions of these methods (e.g. using `caret` package)
3. Teach method to evaluate _any_ statistical method against others.
4. Put all the concepts into a generalized framework of statistical learning
 (that you will promptly find does not work - but hey we tried).

## What this class won't force upon you but you will have to learn anyway

1. Teach you about the data pipeline
    - get the data
    - clean the data
    - access the data (often in databases)
2. Proper programming principles
    - writing good code
    - reproducibility
    - version control
3. Communicating Results
    - Different Audiences
    - Working on a teams






