---
title: "Sample Homework Submission"
author: "Vitaly Druker"
date: "8/28/2018"
output:
  word_document: default
  html_document: default
subtitle: "Homework 1"
---

# Packages Used 

```{r, message=FALSE, warning=FALSE}
library(ISLR)
library(dplyr)
library(ggplot2)
```

# Question 1

_What is the relationship between age and wage? Support your answer with a plot._

The plot below shows a scatter plot of 3,000 subjects age and wage from the Mid-Atlantic region. A smooth line has been applied to the data using the default generalized additive model in the `ggplot2` library. The data shows that wage appears to increase steadily until about 40 years old when it flattens out. The data then appear to stay mostly flat until about 60 when it begins to decline again. 
While this trend is apparent in the data from the smoothing function - there are still a large number of outliers in the top section of the graph that may be skewing the results. 

```{r, message=FALSE}
Wage %>% 
  ggplot(aes(x = age, y = wage)) + 
  geom_point() + 
  geom_smooth()
```


