---
title: "STAT 716 Homework 3"
author: "Vitaly Druker"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)

```

# Exercise 3 (section 3.7)

X_1 = GPA 20
X_2 = IQ .07
X_3 = Gender (1 for female)  35
X_4 = GPA:IQ .01
X_5 = GPA:Gender -10

~ thousand of dollars

```{r}
q3_a <- function(gpa, iq, gender){
  50 + gpa*20 + iq*.07 + gender*35 + gpa*iq*.01 + gpa*gender*-10
}

q3_a(1, 110, 1)
q3_a(1, 110, 0)


q3_a(4, 110, 1)
q3_a(4, 110, 0)


```

i) is not true for low values of GPA (females may make more when there are lower GPAs)

ii) same as above but the other - Not true

iii) TRUE

iv) FALSE


b)
```{r}

q3_a(4, 110, 1)

```

c)

It's size does not imply any/or lack of evidence - it can be a problem with scaled values (e.g. IQ is very large compared with other values in the data set)


Question 13:


```{r}
set.seed(1)
# a
x <- rnorm(100)
# b
eps <- rnorm(100, mean = 0, sd = .25)
# c
y <- -1 + 0.5*x + eps
# d
plot(x, y)
```

Fairly strong linear relationship

```{r}
# e
mod1 <- lm(y ~ x)
summary(mod1)
```

```{r}
# f 
data.frame(x = x, y = y) %>% 
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_abline(slope = .49973, intercept = -1.009)


```

```{r}
#g 
mod2 <- lm(y ~ x + I(x^2))
summary(mod2)

```



