---
title: "STAT 716 - Class 3 - 2018-09-17"
author: "Vitaly Druker"
date: "9/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


```{r}
df_start <- data.frame(x1 = -50:50)
df_start <- df_start %>% 
  mutate(y_ideal = 3*x1 + 2) 

df_start
```


```{r}
# pick some observations from the population
df_observe <- df_start[sample(1:nrow(df_start), size = 100, replace = TRUE), ]

# add some noise
df_observe <- df_observe %>% 
  mutate(y_obs = y_ideal + rnorm(nrow(.), sd = 20))
df_observe
```

```{r}
linear_model <- lm(y_obs ~ x1, data = df_observe)

df_observe %>% 
  ggplot(aes(x = x1, y = y_obs)) +
  geom_point() +
  geom_abline(slope = 3, intercept = 2, color = "blue") +
  geom_smooth(method = "lm", color = 'red')

```
```{r}
summary(linear_model)
```
1st run
std error intercept 1.9, 9.3, .6
std error x1 .06447, .3014, .02




What happens to the estimates if:
What happens if the variance of the error term goes up or down?
What happens if more data is sampled?


```{r}
linear_model


```
```{r}
summary(linear_model)


```

```{r}
confint(linear_model, level = .95)
```

```{r}
t_stat <- seq(-5,5, length.out = 100)

data.frame(t_stat = seq(-5,5, length.out = 100)) %>% 
  mutate(prob_density = dt(t_stat, df = nrow(df_observe)-2)) %>% 
  ggplot(aes(t_stat, prob_density)) + 
  geom_line()+
  geom_vline(xintercept = c(qt(c(.025,.975), df = nrow(df_observe)-2)),
             linetype = "longdash")


```

```{r}
(1 -pt(.596, nrow(df_observe)-2))*2
```

```{r}
data.frame(t_stat = t_stat) %>% 
  mutate(prob_t = pt(t_stat, nrow(df_observe)-2)) %>% 
  ggplot(aes(x=t_stat, y = prob_t)) + geom_line()
```


# Multiple linear Regression

```{r}
ad_data <- read.csv('Advertising.csv')

```

## Section 3.4 - Marketing Plan

Is there a relationship between advertising sales and budget?

F test hypothesis

```{r}
q1_mod <- lm(sales ~ TV  + radio + newspaper, data  = ad_data)

summary(q1_mod)
```

How strong is relationship? 

Look at R^2

```{r}

summary(q1_mod)

```

How large is effect?

```{r}
confint(q1_mod)
```

How do we predict future sales?


```{r}
newdata <- data.frame(TV  = 100, radio = 30, newspaper = 100)
predict(q1_mod, newdata = newdata,interval = "confidence", level = .95)
```
```{r}
predict(q1_mod, newdata = newdata,interval = "prediction")

```

## Other Considerations

```{r}

library(ISLR)
Carseats %>% 
  glimpse 

lm(Sales ~ Income*Advertising, data = Carseats) %>% 
  summary()

```


